---
output:
  pdf_document: 
    number_sections: true
---
\begin{titlepage}

\center
\textsc{\LARGE University of Waterloo}\\[5cm]
{ \huge \bfseries Final Project Report}\\[2cm]
     \textsc{\Large STAT 331 Fall 2018}\\[1cm]
     
\emph{Group 91:}\\
Gnegyao \textsc{Yuan}(20613017)\\
 \textsc{}()\\


\end{titlepage}

\tableofcontents

\newpage


\section{Summary}
The goal of the STAT 331 final project is to explore the relation of healthy male single-fetus birth weight and some explanatory variables. This report will be divided into 4 main sections:  
Summary, which will cover the main purpose of the report and give a brief explanation of how the project will analyze the data. Two candidate models will be produced in the model selection section by using the pre-fitting data diagnostic and automated model selection. Model diagnostics section will perform an in-depth comparison of the two candidates models by comparing different types of residual plots, leverage and influence measures and cross-validation(rPMSE). In the end, there will be a discussion section basing on the result of the most likely linear model we get from the previous sections to talk about several topics such like: ???what is the most important factors associated with/influencing birth weight????  
After using serial statistical analysis way that we learned in STAT 331 course, we find that

\section{Model Selection}

\subsection{Brief data overview and eliminate useless variable}

```{r,echo=FALSE}
library(mice)
chds_births <- read.csv(file = "chds_births.csv")
summary(chds_births)
```
By view the summary of the data, we notic that there are serverial illegal data. The domain of "marital"(the mother's marital status) is 1 to 5 but it is clearly showing that there exist 0 in "marital", we replace all the 0  to NA since it is not avaliubale data(out of range).   
For the categorical predictor "meth"(The self-reported ethnicity of the mothers)
```{r, include=FALSE}
chds_births$marital[!chds_births$marital %in% c(1:5)] = NA
chds_births$meth[chds_births$meth %in% c(0,1,2,3,4,5)] = 0
chds_births$meth[chds_births$meth %in% c(6)] = 1
chds_births$meth[chds_births$meth %in% c(7)] = 2
chds_births$meth[chds_births$meth %in% c(8)] = 3
chds_births$meth[chds_births$meth %in% c(9)] = 4
chds_births$meth[chds_births$meth %in% c(10)] = 5
chds_births$feth[chds_births$feth %in% c(0,1,2,3,4,5)] = 0
chds_births$feth[chds_births$feth %in% c(6)] = 1
chds_births$feth[chds_births$feth %in% c(7)] = 2
chds_births$feth[chds_births$feth %in% c(8)] = 3
chds_births$feth[chds_births$feth %in% c(9)] = 4
chds_births$feth[chds_births$feth %in% c(10)] = 5
chds_births$meth <- as.factor(chds_births$meth)
chds_births$med <- as.factor(chds_births$med)
chds_births$feth <- as.factor(chds_births$feth)
chds_births$fed <- as.factor(chds_births$fed)
chds_births$marital <- as.factor(chds_births$marital)
chds_births$smoke <- as.factor(chds_births$smoke)
chds_births$time <- as.factor(chds_births$time)
chds_births$number <- as.factor(chds_births$number)
```

Since the domain of marital is 1 to 5 and all have actull meaning, by quickly view the dataset, we notice that there are some data out of the domain, so we replace such data to NA since they are NA data.


### Summary of Data
```{r}
summary(chds_births)
```

### Pair plots
```{r}
pairs(~wt + gestation + parity + mage + mht + mwt + fage + fht + fwt,cex = .05,  data = chds_births)
pairs(~wt + gestation + mage + mwt + fage + fwt,cex = .05,  data = chds_births)
```

### remove NA/Inf
```{r}
births_mice <- mice(chds_births)
births_clean <- complete(births_mice)
summary(births_clean)
anyNA(births_clean)
births_clean
levels(births_clean$time)
levels(births_clean$number)
```
MICE helps to impute missing values that are drawn from a distribution specifically designed for each missing data point. Don't like just simply replace all NA variables by mean of the data, MICE can also include the 'var' in the data prediction, which can help lessen the bias and make the data close to the original.

### Revisiting the category variables

```{r}

levels(births_clean$meth)
levels(births_clean$meth)[1] <- "Caucasian"
levels(births_clean$meth)[3] <- "African-American"
levels(births_clean$meth)[c(2,4,5,6)]<- "others"

levels(births_clean$med)[2] <- "middle-school"
levels(births_clean$med)[3] <- "high-school"
levels(births_clean$med)[5] <- "high-school + some college"
levels(births_clean$med)[6] <- "college graduate"
levels(births_clean$med)[c(1,4,7,8)]<- "others"


levels(births_clean$feth)
levels(births_clean$feth)[1] <- "Caucasian"
levels(births_clean$feth)[3] <- "African-American"
levels(births_clean$feth)[c(2,4,5,6)]<- "others"

levels(births_clean$fed)[2] <- "middle-school"
levels(births_clean$fed)[3] <- "high-school"
levels(births_clean$fed)[5] <- "high-school + some college"
levels(births_clean$fed)[6] <- "college graduate"
levels(births_clean$fed)[c(1,4,7,8)]<- "others"


levels(births_clean$marital)[1] <- "married"
levels(births_clean$marital)[c(2,3,4,5)] <- "others"

births_clean$smoke = NULL

levels(births_clean$time)[1] <- "never smoked"
levels(births_clean$time)[2] <- "still smokes"
levels(births_clean$time)[c(3,4,5,6,7,8,9,10)] <- "others"


levels(births_clean$number)[1] <- "never smoked"
levels(births_clean$number)[2] <- "1-4"
levels(births_clean$number)[3] <- "5-9"
levels(births_clean$number)[6] <- "20-29"
levels(births_clean$number)[c(4,5,7,8,9,10)] <- "others"


summary(births_clean)
```



```{r}
M0 <- lm(wt ~ 1, data = births_clean)
Mmax <- lm(wt ~ (. -marital -fed -feth -number -time - meth -med)^2 + marital + fed +feth +number +time + meth + med , data = births_clean)
Mstart <- lm(wt ~ ., data = births_clean)
beta.max <- coef(Mmax)
names(beta.max)[is.na(beta.max)]
anyNA(coef(Mmax))
```


### Forward Model Selection
```{r,incude=False}
Mfwd <- step(object = M0,
               scope = list(lower = M0, upper = Mmax),
               direction = "forward", trace = FALSE)
```

```{r,incude=False}
Mback <- step(object = Mmax, 
              scope = list(lower = M0, upper = Mmax),
              direction = "backward", trace = FALSE)
```

```{r,incude=False}
Mstep <- step(object = Mstart,
              scope = list(lower = M0, upper = Mmax),
              direction = "both", trace = FALSE)
```

```{r}
c(fwd = length(coef(Mfwd)), back = length(coef(Mback)), step = length(coef(Mstep)))
```

## qqplot
```{r}
par(mfrow=c(2,3))
plot(fitted(Mfwd), residuals(Mfwd), main="forward")
plot(fitted(Mback), residuals(Mback), main="backward")
plot(fitted(Mstep), residuals(Mstep), main="stepwise")
qqnorm(residuals(Mfwd), main="forward")
qqnorm(residuals(Mback), main="backward")
qqnorm(residuals(Mstep), main="stepwise")
```
From the Residual vs Fitted plot, it is clearly showing that points are randomly distributed around the 0 line. And the points in QQ-plot almost line on a diagonal line. Both of the two graphs prove that our 

### Press AIC and $R^2$
```{r}
M1 <- Mfwd
M2 <- Mback
M3 <- Mstep
Mnames <- expression(M[FWD], M[BACK], M[STEP])

# press for 3 automated models
press1 <- resid(M1)/(1-hatvalues(M1))
press2 <- resid(M2)/(1-hatvalues(M2))
press3 <- resid(M3)/(1-hatvalues(M3))
PRESS = c(sum(press1^2), sum(press2^2), sum(press3^2))

# R^2 for 3 automated models
r_square1 <- summary(Mfwd)$r.squared
r_square2 <- summary(Mback)$r.squared
r_square3 <- summary(Mstep)$r.squared
R_Squared <- c(r_square1,r_square2,r_square3)

# AIC for 3 automated models
AIC1 <- AIC(M1)
AIC2 <- AIC(M2)
AIC3 <- AIC(M3)
AIC = c(AIC1,AIC2,AIC3)

# display results
disp <- rbind(AIC,PRESS,R_Squared)
colnames(disp) <- Mnames
disp

#plot PRESS statistics
boxplot(x = list(abs(press1),abs(press2),abs(press3)), names = Mnames,
        ylab = expression("|", PRESS[i], "|"),col = c("yellow","orange","violet"))
              
```

### Manual Model
```{r}
Mman1 <- lm(formula = wt ~ gestation + parity + mage + mht + mwt + fage + 
    fht + fwt + income  + number + meth,  data = births_clean)
Mman2 <- lm(formula = wt ~ gestation  + mht + meth + parity + number + 
    mwt + fht + income,  data = births_clean)
Mman3 <- lm(formula = wt ~ gestation + mage + fage  + time + number + meth + feth, 
            data = births_clean)
Mnames <- expression(Mman1, Mman2, Mman3)

# press for 3 automated models
press1 <- resid(Mman1)/(1-hatvalues(Mman1))
press2 <- resid(Mman2)/(1-hatvalues(Mman2))
press3 <- resid(Mman3)/(1-hatvalues(Mman3))
PRESS = c(sum(press1^2), sum(press2^2), sum(press3^2))

# R^2 for 3 automated models
r_square1 <- summary(Mman1)$r.squared
r_square2 <- summary(Mman2)$r.squared
r_square3 <- summary(Mman3)$r.squared
R_Squared <- c(r_square1,r_square2,r_square3)

# AIC for 3 automated models
AIC1 <- AIC(Mman1)
AIC2 <- AIC(Mman2)
AIC3 <- AIC(Mman3)
AIC = c(AIC1,AIC2,AIC3)

# display results
disp <- rbind(AIC,PRESS,R_Squared)
colnames(disp) <- Mnames
disp

#plot PRESS statistics
boxplot(x = list(abs(press1),abs(press2),abs(press3)), names = Mnames,
        ylab = expression("|", PRESS[i], "|"),col = c("yellow","orange","violet"))

```

### model diagnostics
```{r}
Model1 <- Mback
Model2 <- Mman2
```


## Residuals studentlized residuals and standlized residuals
```{r}
Re1 <- residuals(Model1)
Re2 <- residuals(Model2)

StanRe1 <- Re1/sigma(Model1)
StanRe2 <- Re2/sigma(Model2)

StudRe1 <- StanRe1 / sqrt(1-hatvalues(Model1))
StudRe2 <- StanRe2 / sqrt(1-hatvalues(Model2))

```

## Residual plots
```{r}
plot(predict(Model1), Re1, xlab = "Predict Values", ylab = "Residuals", cex.axis = .8,
     abline(h = mean(Re1), col = "red"))
plot(predict(Model2), Re2, xlab = "Predict Values", ylab = "Residuals", cex.axis = .8,
     abline(h = mean(Re2), col = "red"))
```

## Standlized residual plots
```{r}
plot(predict(Model1), StanRe1, xlab = "Predict Values", ylab = "Standlized Residuals", cex.axis = .8,
     abline(h = mean(StanRe1), col = "red"))
plot(predict(Model2), StanRe2, xlab = "Predict Values", ylab = "Standlized Residuals", cex.axis = .8,
     abline(h = mean(StanRe2), col = "red"))
```

## Studentlized residuals plots
```{r}
plot(predict(Model1), StudRe1, xlab = "Predict Values", ylab = "Studentlized Residuals", cex.axis = .8,
     abline(h = mean(StudRe1), col = "red"))
plot(predict(Model2), StudRe2, xlab = "Predict Values", ylab = "Studentlized Residuals", cex.axis = .8,
     abline(h = mean(StudRe2), col = "red"))
```

## PRESS Residuals 
```{r}
press_model1 <- Re1/(1 - hatvalues(Model1))
press_model2 <- Re2/(1 - hatvalues(Model2))
```

## DFFITS Residuals
```{r}
dfts1 <- dffits(Model1)
dfts2 <- dffits(Model2)
```

## Comparison of different residual plots
```{r}
# standlize each of these
p1 <- length(coef(Model1))
n1 <- nobs(Model1)
hbar1 <- p1/n1
StudRe1.stan <- StudRe1 * sqrt(1-hbar1)
press_model1.stan <- press_model1*(1-hbar1)/sigma(Model1)
dfts1.stan <- dfts1*(1-hbar1)/sqrt(hbar1)

# plots all 
par(mfrow = c(1,2), mar = c(4,4,.1,.1))
plot(predict(Model1), rep(0, length(predict(Model1))),
     type = "n",
     ylim = range(StanRe1,StudRe1.stan,dfts1.stan,press_model1.stan),
     xlab = "Predict Values", 
     ylab = "Residuals", 
     cex.axis = .8)
points(predict(Model1), StanRe1, pch = 21, bg = "violet", cex = .8)
points(predict(Model1), StudRe1.stan, pch = 21, bg = "green", cex = .8)
points(predict(Model1), press_model1.stan, pch = 21, bg = "blue", cex = .8)
points(predict(Model1), dfts1.stan, pch = 21, bg = "red", cex = .8)

# standlize each of these
p2 <- length(coef(Model2))
n2 <- nobs(Model2)
hbar2 <- p2/n2
StudRe2.stan <- StudRe2 * sqrt(1-hbar2)
press_model2.stan <- press_model2*(1-hbar2)/sigma(Model2)
dfts2.stan <- dfts2*(1-hbar2)/sqrt(hbar2)

# plots all 
par(mfrow = c(1,2), mar = c(4,4,.1,.1))
plot(predict(Model2), rep(0, length(predict(Model2))),
     type = "n",
     ylim = range(StanRe2,StudRe2.stan,dfts2.stan,press_model2.stan),
     xlab = "Predict Values", 
     ylab = "Residuals", 
     cex.axis = .8)
points(predict(Model2), StanRe2, pch = 21, bg = "violet", cex = .8)
points(predict(Model2), StudRe2.stan, pch = 21, bg = "green", cex = .8)
points(predict(Model2), press_model2.stan, pch = 21, bg = "blue", cex = .8)
points(predict(Model2), dfts2.stan, pch = 21, bg = "red", cex = .8)
     
     
```

## Leverage and influence measures
```{r}
# compute leverage

h1 <- hatvalues(Model1)
h2 <- hatvalues(Model2)

D1 <- cooks.distance(Model1)
D2 <- cooks.distance(Model2)

infl1.ind <- which.max(D1)
infl2.ind <- which.max(D2)

lev1.ind <- h1 > 2*hbar1
lev2.ind <- h2 > 2*hbar2

clrs <- rep("black", len = nobs(Model1))
clrs[lev1.ind] <- "blue"
clrs[infl1.ind] <-"red"
par(mfrow = c(1,1), mar = c(4,4,1,1))
cex <- .8
plot(h1,D1,xlab = "Leverage", ylab = "Cook's Influence Measure", pch = 21, bg = clrs, cex = cex, cex.axis = cex)
abline(v = 2*hbar1, col = "grey60", lty = 2)
legend("topleft", legend = c("High Leverage", "High Influence"), pch = 21, pt.bg = c("blue", "red"),cex = cex, pt.cex = cex)

clrs <- rep("black", len = nobs(Model1))
clrs[lev2.ind] <- "blue"
clrs[infl2.ind] <-"red"
par(mfrow = c(1,1), mar = c(4,4,1,1))
cex <- .8
plot(h2,D2,xlab = "Leverage", ylab = "Cook's Influence Measure", pch = 21, bg = clrs, cex = cex, cex.axis = cex)
abline(v = 2*hbar2, col = "grey60", lty = 2)
legend("topleft", legend = c("High Leverage", "High Influence"), pch = 21, pt.bg = c("blue", "red"),cex = cex, pt.cex = cex)

```


## Residual Histogram
```{r}
par(mfrow=c(1,3))
histogram(residuals(Mfwd)/sigma(Mfwd),breaks = 50,freq = FALSE)
histogram(residuals(Mback)/sigma(Mback),breaks = 50,freq = FALSE)
histogram(residuals(Mstep)/sigma(Mstep),breaks = 50,freq = FALSE)

```




