---
title: "a4p1"
output: pdf_document
---

\subsection{Part a}
```{r setup, include=FALSE}
dataQ1 =matrix(NA, nrow=3, ncol=10)
dataQ1[1,] =c(100, 95, 125, 105, 100, 90, 135, 120, 85, 101)
dataQ1[2,] =c(90, 110, 85, 90, 95, 110, 115, 110, 105, 120)
dataQ1[3,] =c(250, 175, 140, 200, 195, 165, 145, 180, 210, 180)
rownames(dataQ1) =c("Goose Bar", "$9 Dinner", "Carbwich")
```


```{r, echo = FALSE}
boxplot(t(dataQ1),xlab='Group Name',ylab='Contribution')

```
  
The above graph is the  boxplot of the data for all 3 groups.
  
code:  
boxplot(t(dataQ1),xlab='Group Name',ylab='Contribution')  



\subsection{Part b}

Based on the boxplot from part (a), the heights of boxes are a bit different, but overall we can call it seems to be similar variation in each treatment group.

\subsection{Part c}
```{r, echo = FALSE}
contribution = as.numeric(t(dataQ1))
treatments = c(rep('Goose Bar',10), rep('$9 Dinner', 10),rep('Carbwich', 10))
summary( aov(contribution~treatments))
```  
code:  
contribution = as.numeric(t(dataQ1))  
treatments = c(rep('Goose Bar',10), rep('$9 Dinner', 10),rep('Carbwich', 10))  
summary( aov(contribution~treatments))  

\subsection{Part d}
 the p-value is 
$$
\begin{array}{rcl}
3.09*10^{-9}
\end{array}
$$  
code:  
\\begin{array}{rcl}  
3.09*10^{-9}  
\\end{array}  

\subsection{Part e}
$$
\begin{array}{rcl}
f & = & \frac{MS_{TRT}}
              {MS_{RES}}
         \sim F_{dfTRT,dfRES}
\end{array}
$$  
the df TRT is 2 and df RES is 27 from ANOVA table in part(c)
the distribution is   
$$
\begin{array}{rcl}
F_{2,27}
\end{array}
$$
\subsection{Part f}
Since the p-value from part(d) is really small, we reject the null hypothesis that all three approaches lead to the same contribution amounts.  
code:  
\\begin{array}{rcl}  
f & = & \\frac{MS_{TRT}}  
              {MS_{RES}}  
         \\sim F_{dfTRT,dfRES}  
\\end{array}  
\\subsection{Part f}  
\\begin{array}{rcl}  
F_{2,27}  
\\end{array}  


\subsection{Part g}
```{r, echo = FALSE}
mu_hat = mean(dataQ1)
tau_hat = apply(dataQ1,1,mean)
sigmasq_hat = 480
```  
(i) H0: tu1=tu2  
p-value is :
```{r, echo = FALSE}
a = c(1,-1,0)
theta = sum(a * tau_hat)
se_theta = sqrt(sigmasq_hat * sum(a^2)/10)
tobs = theta / se_theta
round(2 * pt(abs(tobs), df = 27, lower.tail = FALSE),3)

```  
code:  
mu_hat = mean(dataQ1)  
tau_hat = apply(dataQ1,1,mean)  
sigmasq_hat = 480  
a = c(1,-1,0)  
theta = sum(a * tau_hat)  
se_theta = sqrt(sigmasq_hat * sum(a^2)/10)  
tobs = theta / se_theta  
round(2 * pt(abs(tobs), df = 27, lower.tail = FALSE),3)  

(ii)  H0:tu1=tu3  
p-value is :
```{r, echo = FALSE}
a = c(1,0,-1)
theta = sum(a * tau_hat)
se_theta = sqrt(sigmasq_hat * sum(a^2)/10)
tobs = theta / se_theta
round(2 * pt(abs(tobs), df = 27, lower.tail = FALSE),3)

```
code:  
mu_hat = mean(dataQ1)  
tau_hat = apply(dataQ1,1,mean)  
sigmasq_hat = 480  
a = c(1,0,-1)  
theta = sum(a * tau_hat)  
se_theta = sqrt(sigmasq_hat * sum(a^2)/10)  
tobs = theta / se_theta  
round(2 * pt(abs(tobs), df = 27, lower.tail = FALSE),3)  

(iii) H0:tu2=tu3  
p-value is :
```{r, echo = FALSE}
a = c(0,1,-1)
theta = sum(a * tau_hat)
se_theta = sqrt(sigmasq_hat * sum(a^2)/10)
tobs = theta / se_theta
round(2 * pt(abs(tobs), df = 27, lower.tail = FALSE),3)

```
code:  
mu_hat = mean(dataQ1)  
tau_hat = apply(dataQ1,1,mean)  
sigmasq_hat = 480  
a = c(0,1-1)  
theta = sum(a * tau_hat)  
se_theta = sqrt(sigmasq_hat * sum(a^2)/10)  
tobs = theta / se_theta  
round(2 * pt(abs(tobs), df = 27, lower.tail = FALSE),3)  

\subsection{Part h}
The null hypothesis means the average contribution from Goose bar should be equal to the 2 times of sum of the rest 2 groups.