†---
title: "STAT331 Project"
author: "Sijie Jin, Raymond Tan, Yinong Wang"
date: '2018-11-14'
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

# Model Selection

## Pre-fitting data diagnostic

```{r}
library(mice)

# read in data
births.raw = read.csv("chds_births.csv")
```

### Summary of Data

```{r}
head(births.raw, 10)
summary(births.raw)

# Validating marital
births.raw$marital[!births.raw$marital %in% c(1:5)] = NA
```

### Categorize data

```{r}
categoricals = c('meth', 'feth', 'med', 'fed', 'marital', 'income', 'smoke', 'time', 'number')#, 'parity', 'mht', 'fht')

# Categorize meth
meth.categories = c(rep('Caucasian', 6), 'Mexican', 'African-American', 'Asian', 'Mixed', 'Other')
for (i in 1:length(meth.categories)) {
  births.raw$meth[births.raw$meth == i-1] = meth.categories[i]
  births.raw$feth[births.raw$feth == i-1] = meth.categories[i]
}
births.raw$meth = as.factor(births.raw$meth)
births.raw$feth = as.factor(births.raw$feth)

# Categorize med
med.categories = c('elementary school', 'middle school', 'high school', 'high school + trade school', 'high school + some college', 'college graduate', 'trade school', 'high school unclear')
for (i in 1:length(med.categories)) {
  births.raw$med[births.raw$med == i-1] = med.categories[i]
  births.raw$fed[births.raw$fed == i-1] = med.categories[i]
}
births.raw$med = as.factor(births.raw$med)
births.raw$fed = as.factor(births.raw$fed)

# Categorize marital
marital.categories = c('married', 'legally separated', 'divorced', 'widowed', 'never married')
for (i in 1:length(marital.categories)) {
  births.raw$marital[births.raw$marital == i] = marital.categories[i]
}
births.raw$marital = as.factor(births.raw$marital)

# Categorize income
income.categories = c('under 2500', '2500-4999', '5000-7499', '7500-9999', '10000-12499', '12500-14999', '15000-17499', '17500-19999', '20000-22499', 'over 22500')
for (i in 1:length(income.categories)) {
  births.raw$income[births.raw$income == i-1] = income.categories[i]
}
births.raw$income = as.factor(births.raw$income)

# Categorize smoke
smoke.categories = c('never', 'smokes now', 'until pregnancy', 'used to, not anymore')
for (i in 1:length(smoke.categories)) {
  births.raw$smoke[births.raw$smoke == i-1] = smoke.categories[i]
}
births.raw$smoke = as.factor(births.raw$smoke)

# Categorize time
time.categories = c('never smoked', 'still smokes', 'during pregnancy', 'less than a year', '1-2 years', '2-3 years', '3-4 years', '5-9 years', 'more than 10 years', 'quit but don’t know when')
for (i in 1:length(time.categories)) {
  births.raw$time[births.raw$time == i-1] = time.categories[i]
}
births.raw$time = as.factor(births.raw$time)

# Categorize number
number.categories = c('never smoked', '1-4', '5-9', '10-14', '15-19', '20-29', '30-39', '40-60', 'more than 60', 'smoked but don’t know how much')
for (i in 1:length(number.categories)) {
  births.raw$number[births.raw$number == i-1] = number.categories[i]
}
births.raw$number = as.factor(births.raw$number)

births.raw
```

### Pair plots

```{r}
births.raw.cont = births.raw[ , ! names(births.raw) %in% categoricals]
summary(births.raw.cont)
pairs(births.raw.cont, cex = 0.2)
```

### NA's and $\pm$Inf's

```{r}
proportion <- sapply(births.raw, function(x) { sum(is.na(x))*100/nrow(births.raw) })
proportion

# births.reduced <- births.raw.cont[, !names(births.raw) %in% c('fht', 'fwt')]

births.mice <- mice(births.raw, method = "cart", seed = 1)
births.clean <- complete(births.mice)
summary(births.clean)
anyNA(births.clean)
```

## Automated Model Selection

```{r}
M0 <- lm(wt ~ 1, data = births.clean) # initial model
Mfull <- lm(wt ~ (.)^2, data = births.clean) # full model
Mstart <- lm(wt ~ ., data = births.clean) # start model

# beta.max <- coef(Mfull)
# names(beta.max)[is.na(beta.max)]
```

### Forward Model Selection

```{r}
# forward
Mfwd <- step(object = M0,
             scope = list(lower = M0, upper = Mfull),
             direction = "forward",
             trace = FALSE)
```

### Backward Model Selection

```{r}
# backward
Mback <- step(object = Mfull, 
              scope = list(lower = M0, upper = Mfull),
              direction = "backward", 
              trace = FALSE)
```

### Stepwise Model Selection

```{r}
# stepward
Mstep <- step(object = Mstart, 
              scope = list(lower = M0, upper = Mfull),
              direction = "both", 
              trace = FALSE)
```

## Manual Model Selection

### Cross Validation

```{r}

```

### PRESS Statistics

```{r}
#Press Statistic
calPress <- function(M){
  press <- resid(M)/(1-hatvalues(M))
  return(press)
}

press1 <- calPress(M1);
press2 <- calPress(M2);
```



### Akaike Information Criterion
```{r}
# AIC Statistic for M1
AIC1 <- AIC(M1)

#AIC Statistic for M2
AIC2 <- AIC(M2)
```



## Conclusion

# Model Diagnostics

## Linear Model Assumptions
```{r}
M1 <- Mfwd
M2 <- Mstep

summary(M1)
summary(M2)
#anova(M1, M2)
```

```{r}
cex <- .8

# studendized residual for M1:
res1 <- residuals(M1) # Hat matrix
h1 <- hatvalues(M1)
sigma.hat1 <- sigma(M1)
res.stu1 <- res1/sqrt(1-h1) # studentized residuals, but on the data scale
y.hat1 <- predict(M1)
par(mar = c(4,4,.1,.1))
plot(y.hat1, res1, pch = 21, bg = "black", cex = cex, cex.axis = cex,
        xlab = "Predicted Birth Weight in M1",
        ylab = "Residual Birth Weight in M1",
        abline(h = mean(res1), col = "blue"))
points(y.hat1, res.stu1, pch = 21, bg = "red", cex = cex)
legend(x = "bottomleft", c("Residuals", "Studentized Residuals"),
pch = 21, pt.bg = c("black", "red"), pt.cex = cex, cex = cex)


# studendized residual for M2:
res2 <- residuals(M2) 
X2 <- model.matrix(M2) # design matrix
h2 <- hatvalues(M2)
sigma.hat2 <- sigma(M2)
res2 <- residuals(M2)
res.stu2 <- res2/sqrt(1-h2) # studentized residuals, but on the data scale
y.hat2 <- predict(M2)
par(mar = c(4,4,.1,.1))
plot(y.hat2, res2, pch = 21, bg = "black", cex = cex, cex.axis = cex,
    xlab = "Predicted Birth Weight in M2", 
    ylab = "Residual Birth Weight in M2", 
    abline(h = mean(res2), col = "blue"))
points(y.hat2, res.stu2, pch = 21, bg = "red", cex = cex)
legend(x = "bottomleft", c("Residuals", "Studentized Residuals"),
pch = 21, pt.bg = c("black", "red"), pt.cex = cex, cex = cex)


# standardized residual for M1:
par(mar = c(4,4,.1,.1))
hist(res1/sigma.hat1, breaks = 50, freq = FALSE, cex.axis = cex,
xlab = "Standardized Residual for M1", main = "")
curve(dnorm(x), col = "red", add = TRUE)
abline(v = mean(res1/sigma.hat1), col = "green")

# standardized residual for M2
par(mar = c(4,4,.1,.1))
hist(res2/sigma.hat2, breaks = 50, freq = FALSE, cex.axis = cex,
xlab = "Standardized Residual for M2", main = "")
curve(dnorm(x), col = "red", add = TRUE)
abline(v = mean(res2/sigma.hat2), col = "green")

# QQ-plot for M1
qqnorm(res.stu1, ylim = c(-40, 60))
qqline(res.stu1)
# head(res.stu1)

# QQ-plot for M2
qqnorm(res.stu2, ylim = c(-40, 60))
qqline(res.stu2)
# head(res.stu2)
```

## Colinearity

## Outliers

# Conclusions

